---
title: "Exercise 9.4"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

## Load packages

Here is the R code to download the required packages for this exercise.

```{r pac-load, exercise.eval=TRUE}
# install package manager 'pacman'
if (!require(pacman)){
  install.packages('pacman')
}

# load packages needed for this exercise
library(pacman)
p_load(tidyverse,
       lctools, # to calculate Moran's I
       spdep, # to calculate geary's c
       geoR, # to compute variogram
       gridExtra, # to stack plots
       gstat, automap, nlme, # packages for variogram model selection
       sp # need a function called `coordinates`
       )

```

## Data

This is equivalent to data step in SAS. Here, the data is entered inside a function called `tibble`.

```{r data-import, exercise.eval=TRUE}
# Enter data
a <- tibble(I = 1:16, YI = c(41, 60, 81, 22, 8, 20, 28, 2,
                             0, 2, 2, 8, 0, 43, 61, 50)) %>% 
  # creat new variable East and North
  mutate(East = 1,
         North = I)

# print the data
a
```

## Autocorrelation statistics

```{r autocorr-statistics, exercise.eval = TRUE}

# visualize the data
ggplot(data = a) +
  geom_point(mapping = aes(x = East, y = North, size = YI, color = YI)) +
  ggtitle("Spatial Distribution of YI Observation") +
  theme(plot.title = element_text(hjust = 0.5))

# calculate Moran's I
Coords <- a %>% 
  dplyr::select(East, North)

mI <- moransI(Coords, Bandwidth = 1, a$YI)

# print Moran's I table
moran.table <- tribble(
  ~`Moran's I`, ~`Expected I`, ~`Z randomization`, ~`P value randomization`,
  #------------/--------------/-------------------/------------------------
  mI$Morans.I, mI$Expected.I,  mI$z.randomization, mI$p.value.randomization
  )

moran.table

# create Moran's I scatter plot
l.moran <- l.moransI(Coords, Bandwidth = 1, a$YI)

# calculate geary's c
Coords_num <- coordinates(Coords)

# create an object of class 'nb' so that it can be used with function from packege `spdep`
Coords_nb <- knn2nb(knearneigh(Coords_num))

# create a 'listw' object for use in the function `geary.test`
coords_listw <- nb2listw(Coords_nb)

gearyC <- geary.test(a$YI, coords_listw, alternative = "two.sided")
gearyC

```

## First variogram

```{r first-variogram-model, exercise.eval=TRUE}
v1 <- variog(coords = Coords_num, data = a$YI, breaks = seq(0.5, 15.5),
             max.dist = 11)

# extract data from object v1 for plotting
v1_plot_data <- cbind(v1$u, v1$v, v1$n) %>% 
  as.data.frame() %>% 
  dplyr::rename(Distance = V1,
                Semivariance = V2,
                Pair_count = V3)
v1_plot_data

# plot variogram
v1_plot_vario <- ggplot(data = v1_plot_data) +
  geom_point(mapping = aes(x = Distance, y = Semivariance)) +
  ggtitle("Empirical Semivariogram of YI") +
  theme(plot.title = element_text(hjust = 0.5))

# plot pair counts
v1_plot_pair_count <- ggplot(data = v1_plot_data) +
  geom_col(mapping = aes(x = Distance, y = Pair_count), width = 0.01, color = "blue")

# stack two plots
grid.arrange(v1_plot_vario, v1_plot_pair_count,
             ncol = 1, heights = c(3, 1))
```

## Second variogram  

Plot robust and classical variogram together.  

```{r second-variogram-model, exercise.eval = TRUE}
# plot robust and classical variogram on the same plot  

# fit robust variogram
v1_robust <- variog(coords = Coords_num, data = a$YI, breaks = seq(0.5, 15.5),
             max.dist = 11, estimator.type = "modulus")

# extract the data
v1_robust_data <- cbind(v1_robust$u, v1_robust$v, v1_robust$n) %>% 
  as.data.frame() %>% 
  dplyr::rename(Distance = V1,
                Semivariance = V2,
                Pair_count = V3)

# plot robust variogram
v1_robust_vario <- ggplot(data = v1_robust_data) +
  geom_point(mapping = aes(x = Distance, y = Semivariance)) +
  ggtitle("Empirical Semivariogram of YI - Robust estimation") +
  theme(plot.title = element_text(hjust = 0.5))

v1_robust_vario

# combine robust and classical variogram  
var_comb <- v1_robust_data %>% 
  
  # combine robust and classical variogram datasets
  dplyr::rename(Semivariance_robust = Semivariance) %>% 
  bind_cols(dplyr::select(v1_plot_data, Semivariance)) %>% 
  gather(key = "Semivariance_type", value = "Semivariance", -c(Distance, Pair_count)) %>% 
  
  # plot
  ggplot() +
  geom_point(mapping = aes(x = Distance, y = Semivariance, color = Semivariance_type)) +
  ggtitle("Empirical Semivariogram for YI") +
  theme(plot.title = element_text(hjust = 0.5))

var_comb
```  

## Variogram model selection 

We will use the package `gstat` for variogram model selection

```{r model-selection}
# fit the model 
coordinates(a) = ~East+North
v <- variogram(YI ~ East + North, data = a, alpha = c(0, 45, 90, 135))
v
plot(v)
fit.variogram(v, vgm(c("Exp", "Sph", "Gau")))

v_emp <- variogram(YI ~ East + North, data = a, cutoff = 11)
v_emp
plot(v_emp)

v_exp <- fit.variogram(v_emp, vgm("Exp"))
v_exp

plot(v_emp, v_exp, ylim = c(0,1200))

autofitVariogram(YI ~ East + North, a, model = c("Sph", "Exp", "Gau"))

# try nlme package
m <- gls(YI ~ 1, data = a)
m1 <- gls(YI ~ 1, correlation = corExp(form = ~East + North, nugget = TRUE), data = a)

m2 <- gls(YI ~1, correlation = corGaus(form = ~East + North, nugget = TRUE), data = a)

m3 <- gls(YI ~1, correlation = corSpher(form = ~East + North, nugget = TRUE), data = a)

AIC(m1, m2, m3)

variom1 <- Variogram(m1, form = ~East + North, resType = "pearson")
plot(variom1, xlim = c(0,12))

variom2 <- Variogram(m2, form = ~East + North, resType = "pearson")
variom3 <- Variogram(m3, form = ~East + North, resType = "pearson")

```

